Omics-based Single Step Trait Prediction

### 2017-03-28 paper/manuscript.tex
Substitute tables with figures, use r from imputed genos only

\section*{Results}
\Genetics2level{Inbred lines}
At the level of the 149 genotypes for which both, genomic (G) and 
transcriptomic (T) information, were available, the use of T yielded higher
predictive abilities than the use of G for four out of six traits
(Table \ref{table:InbredTable}).
In terms of predictive ability, the absolute advantage of T over G was as large 
as 0.04 as observed for the traits '100grainweight', 'Eardiameter' and 
'Silkingtime'.
Except for 'Kernelwidth', the predictive ability for predictions based on
G and the full set of inbred lines ($n = 211$) was always higher than for the 
reduced set of inbred lines ($n = 149$).
The combination of G and T (\textit{i.e.}, GT) in a single-step approach for the 
prediction of all 211 lines performed at least as well as G alone when T was 
superior over G for the reduced set of 149 lines.
In the case of '100grainweight' and 'Eardiameter', the use of GT yielded a 
slight improvement in predictive ability over G.
Where T was a worse predictor for the reduced set of 149 lines than G,
predictive abilities of GT for the full set of 211 lines were also worse than 
those based on G alone.

% latex table generated in R 3.3.3 by xtable 1.8-2 package
% Fri Mar 24 11:34:47 2017
% Script: ./analysis/maizego_core_fraction_plot_predictive_abilities.R
\begin{table}[ht]
\centering
\caption{Predictive abilities and corresponding coefficients of variation for different core sets from the tropical/subtropical material of the maize diversity panel. As predictors, genomic (G) data, transcriptomic (T) data and their combination (GT) were used. The decimal after the predictor name indicates the size of the genetic material that was covered by inbred lines in the core set.} 
\label{table:InbredTable}
\begin{tabular}{rllllll}
  \toprule
 & 100grainweight & cobweight & Eardiameter & Kernelwidth & Plantheight & Silkingtime \\ 
  \midrule
\multicolumn{6}{l}{{\bfseries Only incomplete predictor: FALSE}}\\
G Full & 0.43 (0.073) & 0.37 (0.173) & 0.39 (0.053) & 0.34 (0.045) & 0.45 (0.076) & 0.53 (0.03) \\ 
  GT Full & 0.44 (0.075) & 0.33 (0.173) & 0.4 (0.054) & 0.34 (0.043) & 0.4 (0.071) & 0.53 (0.028) \\ 
   \midrule
\multicolumn{6}{l}{{\bfseries Only incomplete predictor: TRUE}}\\
G 1.0 & 0.36 (0.077) & 0.21 (0.168) & 0.3 (0.053) & 0.34 (0.047) & 0.41 (0.078) & 0.45 (0.03) \\ 
  T 1.0 & 0.4 (0.075) & 0.15 (0.157) & 0.37 (0.051) & 0.36 (0.042) & 0.32 (0.069) & 0.49 (0.028) \\ 
  GT 0.1 & 0.29 (0.076) & 0.21 (0.166) & 0.23 (0.052) & 0.32 (0.047) & 0.4 (0.074) & 0.39 (0.03) \\ 
  GT 0.2 & 0.35 (0.072) & 0.23 (0.164) & 0.26 (0.052) & 0.36 (0.047) & 0.38 (0.072) & 0.44 (0.027) \\ 
  GT 0.3 & 0.36 (0.071) & 0.2 (0.157) & 0.28 (0.05) & 0.37 (0.044) & 0.42 (0.074) & 0.5 (0.028) \\ 
  GT 0.4 & 0.39 (0.07) & 0.2 (0.156) & 0.33 (0.049) & 0.35 (0.043) & 0.38 (0.073) & 0.52 (0.028) \\ 
  GT 0.5 & 0.41 (0.067) & 0.16 (0.155) & 0.31 (0.049) & 0.4 (0.041) & 0.44 (0.075) & 0.49 (0.028) \\ 
  GT 0.6 & 0.4 (0.074) & 0.13 (0.156) & 0.27 (0.05) & 0.37 (0.042) & 0.42 (0.075) & 0.55 (0.027) \\ 
  GT 0.7 & 0.37 (0.073) & 0.1 (0.156) & 0.31 (0.051) & 0.37 (0.042) & 0.33 (0.075) & 0.52 (0.028) \\ 
  GT 0.8 & 0.4 (0.077) & 0.21 (0.158) & 0.36 (0.051) & 0.41 (0.046) & 0.29 (0.074) & 0.49 (0.029) \\ 
  GT 0.9 & 0.39 (0.079) & 0.21 (0.169) & 0.38 (0.056) & 0.41 (0.046) & 0.29 (0.075) & 0.49 (0.031) \\ 
  GT A1 & 0.32 (0.079) & 0.03 (0.155) & 0.28 (0.053) & 0.32 (0.045) & 0.31 (0.073) & 0.46 (0.03) \\ 
  GT A2 & 0.4 (0.081) & 0.18 (0.164) & 0.32 (0.055) & 0.33 (0.046) & 0.36 (0.073) & 0.47 (0.03) \\ 
  GT A3 & 0.3 (0.071) & 0.1 (0.164) & 0.34 (0.054) & 0.41 (0.041) & 0.33 (0.072) & 0.41 (0.029) \\ 
   \bottomrule
\multicolumn{6}{l}{}\\
\end{tabular}
\end{table}

\Genetics2level{Hybrids}
The maize hybrid data set contains information on six agronomic traits and
three predictors, namely pedigree data (P), genomic information (G), 
transcriptomic information (T), and combinations thereof.
Pedigree records and genomic information were available for the 245 parent 
lines of all 1,521 hybrids and we denote these data as the 'Full' set whereas
transcriptomic data were available for 103 parent lines of 685 hybrids, which
we denote as 'Core'.
For dry matter yield (DMY), T was the best predictor in the core set of 685 
hybrids with a predictive ability that was 0.08 points higher than for genomic 
data, as the second best predictor for this trait 
(Table \ref{table:HybridTable}).
In the core set, P had the highest predictive ability for starch ('STA').
For the other four traits, G had the best predictive abilities.
In the full set of 1,521 hybrids, predictive abilities obtained with G rose 
markedly for all traits with the exception of 'FAT'.
The combination of genomic and transriptomic information (GT) in a single-step
model yielded a small improvement of G alone for the trait 'DMY'.
All traits, where T was not superior to G in the core set, did not benefit from 
the single-step approach when compared to genomic information alone.
Predictive abilities resulting from combinations of pedigree with genomic (PG)
information were better than P alone for all traits and PT improved upon P in
the core set for four traits.

% latex table generated in R 3.3.3 by xtable 1.8-2 package
% Fri Mar 24 12:24:30 2017
% Script: ./analysis/uhoh_plot_predictive_abilities.R
\begin{table}[ht]
\centering
\caption{Predictive abilities and corresponding coefficients of variation for the set of maize hybrids. As predictors, pedigree (P), genomic (G), transcriptomic (T) data and combinations thereof we used.} 
\label{table:HybridTable}
\begin{tabular}{lllllll}
  \toprule
Predictor & DMY & DMC & FAT & PRO & STA & SUG \\ 
  \midrule
P Full & 0.54 (0.049) & 0.48 (0.052) & 0.21 (0.037) & 0.6 (0.019) & 0.35 (0.043) & 0.42 (0.117) \\ 
  G Full & 0.74 (0.052) & 0.63 (0.052) & 0.43 (0.036) & 0.65 (0.022) & 0.46 (0.046) & 0.47 (0.126) \\ 
  PG Full & 0.71 (0.054) & 0.57 (0.055) & 0.35 (0.038) & 0.62 (0.023) & 0.39 (0.048) & 0.42 (0.13) \\ 
  PT Full & 0.73 (0.056) & 0.53 (0.059) & 0.35 (0.04) & 0.61 (0.025) & 0.31 (0.05) & 0.37 (0.14) \\ 
  GT Full & 0.75 (0.053) & 0.61 (0.055) & 0.43 (0.037) & 0.65 (0.024) & 0.44 (0.048) & 0.44 (0.134) \\ 
  P Core & 0.46 (0.047) & 0.53 (0.05) & 0.23 (0.037) & 0.46 (0.023) & 0.38 (0.039) & 0.34 (0.132) \\ 
  G Core & 0.67 (0.053) & 0.47 (0.055) & 0.43 (0.038) & 0.48 (0.027) & 0.36 (0.046) & 0.35 (0.132) \\ 
  T Core & 0.75 (0.055) & 0.34 (0.062) & 0.4 (0.041) & 0.47 (0.029) & 0.16 (0.05) & 0.27 (0.145) \\ 
   \bottomrule
\end{tabular}
\end{table}

@@

Predictive abilities obtained with the different core sets varied widely within
traits (Fig. \ref{fig:CoreSetInfluence}, Table \ref{table:InbredTable}).
For the traits 'cobweight' and 'Plantheight', which were better predicted via G
than via T in the reduced set of lines, predictive abilities were higher when
only a small number of genotypes was included in the core set
For most other traits, an initial increase in predictive abilities when moving
from a core set size of 0.1 to 0.5/0.6 was observed, followed by a drop around
0.6/0.7 and a subsequent hike up to a core set size of 0.9.

\begin{figure}[H]
\centering
\includegraphics{./tables_figures/core_fraction_predictive_ability_trend.pdf}
\caption{
Influence of core set composition on predictive abilities.
Predictive abilities are based on single step predition with genomic data as the 
complete predictor and transcriptomic data as the incomplete predictor.
Results are shown for six agronomic traits and nine core sets of the maize 
diversity panel.
}
\label{fig:CoreSetInfluence}
\end{figure}

@@

Of the 149 genotypes, 17 belonged to more than 50\% to the first ancestral
population (A1), 109 belonged to more than 50\% to the second ancestral
population (A2) and 19 belonged to more than 50\% to the third ancestral
population while four genotypes could not be unequivocally assigned to either
ancestral population (mixed).
To adjust the size of the set of genotypes largely belonging to A2 to the other
two sets, we selected 19 genotypes from A2 using previously described core
sampling.
For single-step prediction, each of the three genotype sets were then declared
as having genomic and transcriptomic information whereas their complements were
covered by transcriptomic data only.
Genotypes from A1 and A3, which were at the far ends in the PCA plot (Fig.
\ref{fig:PopStructure}B), were estimated with less precision than genotypes
from the central cluster of genotypes belonging largely to the A2 population
(Fig. \ref{fig:PopStructure}A, Table \ref{table:InbredTable}).

>>>>

\section*{Results}
\Genetics2level{Inbred lines}
At the level of the 149 genotypes for which both, genomic (G) and 
transcriptomic (T) information, were available, the use of T yielded higher
predictive abilities than the use of G for four out of six traits
(Table \ref{table:InbredResults}).
In terms of predictive ability, the absolute advantage of T over G was as large 
as 0.04 as observed for the traits '100grainweight', 'Eardiameter' and 
'Silkingtime'.
Except for 'Kernelwidth', the predictive ability for predictions based on
G and the full set of inbred lines ($n = 211$) was always higher than for the 
reduced set of inbred lines ($n = 149$).
The combination of G and T (\textit{i.e.}, GT) in a single-step approach for the 
prediction of all 211 lines performed at least as well as G alone when T was 
superior over G for the reduced set of 149 lines.
In the case of '100grainweight' and 'Eardiameter', the use of GT yielded a 
slight improvement in predictive ability over G.
Where T was a worse predictor for the reduced set of 149 lines than G,
predictive abilities of GT for the full set of 211 lines were also worse than 
those based on G alone.

% Script: ./analysis/maizego_core_fraction_plot_predictive_abilities.R
\begin{figure}[H]
  \centering
  \includegraphics{./tables_figures/inbred_line_combi_plot.pdf}
  \caption{
  Predictive abilities and corresponding coefficients of variation for 
  different core sets from the tropical/subtropical material of the maize 
  diversity panel.
  As predictors, genomic (G) data, transcriptomic (T) data and their 
  combination (GT) were used.
  The decimal after the predictor name indicates the size of the genetic 
  material that was covered by inbred lines in the core set.
  Influence of core set composition on predictive abilities.
  Predictive abilities are based on single step predition with genomic data as the 
  complete predictor and transcriptomic data as the incomplete predictor.
  Results are shown for six agronomic traits and nine core sets of the maize 
  diversity panel.
  } 
  \label{fig:InbredResults}
\end{figure}

\Genetics2level{Hybrids}
The maize hybrid data set contains information on six agronomic traits and
three predictors, namely pedigree data (P), genomic information (G), 
transcriptomic information (T), and combinations thereof.
Pedigree records and genomic information were available for the 245 parent 
lines of all 1,521 hybrids and we denote these data as the 'Full' set whereas
transcriptomic data were available for 103 parent lines of 685 hybrids, which
we denote as 'Core'.
For dry matter yield (DMY), T was the best predictor in the core set of 685 
hybrids with a predictive ability that was 0.08 points higher than for genomic 
data, as the second best predictor for this trait 
(Table \ref{fig:HybridResults}).
In the core set, P had the highest predictive ability for starch ('STA').
For the other four traits, G had the best predictive abilities.
In the full set of 1,521 hybrids, predictive abilities obtained with G rose 
markedly for all traits with the exception of 'FAT'.
The combination of genomic and transriptomic information (GT) in a single-step
model yielded a small improvement of G alone for the trait 'DMY'.
All traits, where T was not superior to G in the core set, did not benefit from 
the single-step approach when compared to genomic information alone.
Predictive abilities resulting from combinations of pedigree with genomic (PG)
information were better than P alone for all traits and PT improved upon P in
the core set for four traits.

% latex table generated in R 3.3.3 by xtable 1.8-2 package
% Fri Mar 24 12:24:30 2017
% Script: ./analysis/uhoh_plot_predictive_abilities.R
\begin{figure}[H]
\centering
  \includegraphics{./tables_figures/hybrid_predictive_ability.pdf}
  \caption{
    Predictive abilities and corresponding coefficients of variation for the 
    set of maize hybrids.
    As predictors, pedigree (P), genomic (G), transcriptomic (T) data and 
    combinations thereof we used.
  } 
  \label{fig:HybridResults}
\end{figure}

@@

Predictive abilities obtained with the different core sets varied widely within
traits (Fig. \ref{fig:InbredResults}C).
For the traits 'cobweight' and 'Plantheight', which were better predicted via G
than via T in the reduced set of lines, predictive abilities were higher when
only a small number of genotypes was included in the core set
For most other traits, an initial increase in predictive abilities when moving
from a core set size of 0.1 to 0.5/0.6 was observed, followed by a drop around
0.6/0.7 and a subsequent hike up to a core set size of 0.9.

@@

Of the 149 genotypes, 17 belonged to more than 50\% to the first ancestral
population (A1), 109 belonged to more than 50\% to the second ancestral
population (A2) and 19 belonged to more than 50\% to the third ancestral
population while four genotypes could not be unequivocally assigned to either
ancestral population (mixed).
To adjust the size of the set of genotypes largely belonging to A2 to the other
two sets, we selected 19 genotypes from A2 using previously described core
sampling.
For single-step prediction, each of the three genotype sets were then declared
as having genomic and transcriptomic information whereas their complements were
covered by transcriptomic data only.
Genotypes from A1 and A3, which were at the far ends in the PCA plot (Fig.
\ref{fig:PopStructure}B), were estimated with less precision than genotypes
from the central cluster of genotypes belonging largely to the A2 population
(Fig. \ref{fig:PopStructure}A, Table \ref{fig:InbredTable}).
